Clustering analysis and statistics

### 2019-08-20 notebooks/Clustering analysis.Rmd
removed cgmlst results and added evolutionary models to the trees

```{r libraries, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggtree)
library(ape)
library(tibble)
library(knitr)
library(vegan)
library(distanceR)
library(impoRt)
library(readxl)
library(readr)
library(iteRate)
library(treeio)
library(adephylo)
library(phangorn)
library(RColorBrewer)
library(kableExtra)
```

@@

# Minimum spanning tree
The EnteroBase cgMLST scheme [1] for _E. coli_ was used with chewBBACA [2] to generate the cgMLST data.
A minimum spanning tree calculated from the cgMLST data with GrapeTree [3] using MSTreeV2. Number of allele differences are denoted as numbers on branches between nodes. The branches have been collapsed on ten differences or less. The number inside each node is the sequence type based on the EnteroBase 7-gene MLST scheme [4].
```{r, fig.width=15, fig.height=12}
include_graphics("../figures/GrapeTree_cgMLST.png")
```

@@

# Maximum likelihood trees on clades of interest
Each tree is midpoint rooted for visual purposes. The values on each node is the UF Bootstrap values calculated by IQTree. The tip point shape represent the animal species of origin, and the color the year of isolation. The tip labels represent the location of the isolate; the first number represent the county, and the second number the municipality. 
```{r message=FALSE, warning=FALSE}
# Import isolate information and tree metadata
location_data <- read_excel("../data/st_location_data.xlsx") %>%
  select(id, Kommune, Fylke) %>%
  mutate(fyl_id = group_indices(., Fylke),
         kom_id = group_indices(., Kommune)) %>%
  mutate(Location = paste(fyl_id, kom_id, sep = "-")) %>%
  select(id, Location)

@@

# References
[1] Alikhan N-F et al. A genomic overview of the population structure of Salmonella. Casadesús J, editor. PLOS Genet [Internet]. 2018 Apr 5;14(4):e1007261. Available from: https://dx.plos.org/10.1371/journal.pgen.1007261

[2] Machado MP et al. chewBBACA: A complete suite for gene-by-gene schema creation and strain identification. Microb Genomics. 2018;1–7.

[3] Z Zhou et al. (2018) “GrapeTree: Visualization of core genomic relationships among 100,000 bacterial pathogens”, Genome Res. doi: https://doi.org/10.1101/gr.232397.117

[4] Wirth T et al. (2006) "Sex and virulence in Escherichia coli : an evolutionary perspective", Molecular Microbiology (2006) 60 (5), 1136–1151, doi: 10.1111/j.1365-2958.2006.05172.x

>>>>

```{r libraries, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggtree)
library(ape)
library(tibble)
library(vegan)
library(distanceR)
library(impoRt)
library(readxl)
library(readr)
library(iteRate)
library(treeio)
library(adephylo)
library(phangorn)
library(RColorBrewer)
library(kableExtra)
```

@@

@@

# Maximum likelihood trees on clades of interest
Each tree is midpoint rooted for visual purposes. The values on each node is the UF Bootstrap values calculated by IQTree. The tip point shape represent the animal species of origin, and the color the year of isolation. The tip labels represent the location of the isolate; the first number represent the county, and the second number the municipality.

Evolutionary models:

- Clade A: K2P+ASC+R2

- Clade C: K3P+ASC

- Clade D: K2P+ASC+R2

- Clade E: K3P+ASC

```{r message=FALSE, warning=FALSE}
# Import isolate information and tree metadata
location_data <- read_excel("../data/st_location_data.xlsx") %>%
  select(id, Kommune, Fylke) %>%
  mutate(fyl_id = group_indices(., Fylke),
         kom_id = group_indices(., Kommune)) %>%
  mutate(Location = paste(fyl_id, kom_id, sep = "-")) %>%
  select(id, Location)

@@
